Flattering Advice: Avoiding Disappointment in Advice-Giving

### 2025-01-28 manuscript/manuscript.qmd
Revised version :)

---
title: "Flattering Advice: Avoiding Disappointment in Advice-Giving"
shorttitle: "Flattering Advice"
date: yes
author: 
  - name: "Amanda Chen"
    corresponding: true
    orcid: 0000-0002-4428-627X
    email: zchengj@connect.ust.hk
    affiliations:
      - id: id1
        name: "The Hong Kong University of Science and Technology"
        department: "Department of Management"
        city: "Hong Kong"
  - name: "David Hagmann"
    orcid: 0000-0002-2080-997X
    affiliations:
    - id: id1
author-note:
  disclosures:
    gratitude: The authors are grateful to XXX for their helpful comments.
abstract: Good advice improves decision quality but often requires delivering unpleasant truths that may disappoint advisees. Across three pre-registered and incentivized experiments involving real adviser-advisee interactions (N = 3,900), we show that advisers prioritize avoiding disappointment at the expense of accuracy and their own earnings. In Study 1, advisers financially rewarded for accuracy still tailor recommendations to aspirational goals expressed by advisees, resulting in worse advice. When incentivized to be liked, advisers provide even more flattering advice, and advisees reward this by rating these advisers as more likable, despite the advice being less honest and less accurate (Study 2). The desire to avoid disappointment may lead to inequities if there are differences in expectations across social groups. In Study 3, we examine a setting in which men expect to perform better than women. We show that advisers take into account these expectations, leading to systematically different advice for men and women even when their gender is unknown to advisers. Advisers' efforts to avoid disappointment may thus contribute to systematic gender disparities in advice, with implications for downstream decision-making.

@@

We present the results of three preregistered experiments in which
participants are paired anonymously as advisers and advisees. Advisers
are incentivized to offer accurate advice, with bonus earnings depending
on the outcome of an advisee's decision. We study a setting in which
advisees can communicate their expectations to an adviser, and
experimentally manipulate whether advisers observe this expectation. Our
design simulates organizational contexts where mentors and managers
often have some sense of an employee's confidence or aspirations and
thus could adjust their feedback to avoid disappointing them.
Interpersonal considerations would likely be stronger in situations in
which the adviser and advisee have an existing relationship and when the
advice is delivered face-to-face. Moreover, advisers usually do not
suffer any direct costs when their flattering advice leads to a poor
outcome for the recipient, particularly when it is not clear what would
have happened under a counterfactual.\^\[In some cases, however,
repeated interaction might also increase incentives for honest feedback.
Someone who is known to persistently give overoptimistic advice may be
viewed as less trustworthy in the long run.\]

@@

```{r}
#| label: tbl-study1regs
#| tbl-cap: "Advice to compete against the High Performer Group in Study 1. Displaying the non-binding choice of those for whom the Low Performer Group was selected by default makes it less likely that they are advised to compete against the High Performer Group (Column 1). Column 2 restricts the analysis to advisers who observe advisees' initial non-binding choice and controls for advisees' score on the quiz. Standard errors are clustered at the adviser level."

@@

We now examine whether advisers take the advisees' expectations into
account. Advisers observe the true performance of the advisee, such that
expectations about performance add no additional information. However,
because participants who score higher are also more likely to pick the
High Performer Group, we use our Default treatment as an instrument that reflect
higher expectations.
We show linear probability models for the decision of the Advisers, which is
equal to 1 if they recommend competing against the High Performer Group and
0 if they recommend the Low Performer Group. In Column 1 of @tbl-study1regs,
we show our main specificiation with the experimental assignment for Advisees
(High Performer Default vs. Low Performer Default) and Advisers
(Expectation Shown vs. Expectation Hidden), as well as their interaction.
Because each adviser made ten
recommendations, we cluster standard errors at the adviser level. We find that
displaying expectations of those in the Low Performer Default reduces recommendations
to compete against the Top group by 4 percentage points, or about 10\% relative
to when expectations are hidden. This decrease is marginally smaller for advisees
with a default choice to compete against the High Performer Group
(nudged to have high expectations, though not necessarily holding them), and whose
advice did not differ based on whether their expectations were shown or hidden.
We reported a linear probability model on advising to
compete against the High Performer group for advisers in the
"Expectations" treatment, controlling for the true performance of the
advisee in Column 2 of @tbl-study1regs. We observe a significant main
effect of the default treatment, where a high default increased the
likelihood of receiving flattering advice. It is worth noting that,
since only
`r round(mean(study1_time1[study1_time1$default == "High",]$Guess),2)*100`
% of advisees in the high default group actually stuck with their
default choice, this represents a conservative test of the effect of
high expectations on the likelihood of receiving flattering advice.

@@

Next, we recruit `r scales::comma(dim(study2_time2)[1])` participants from Prolific
for the role of advisers ($M_{\text{Age}}$ = `r mean(study2_time2$age)` years;
`r round(mean(study2_time2$gender == "Female")*100, 2)`% Female). We first
collect demographic information, then match them to a group of
the opposite sex. They rank selfies from ten participants from most to least
attractive using a drop-down menu next to each picture. Because of a
limitation with the survey software, we could not validate that each rank is
given only once, and we remove 115 participants who failed to follow instructions 
and did not provide a complete ranking.^[Moreover, we could not collect data from any participant who did not select any participant as the 7th most attractive participant, as that image would be shown on subsequent pages.]

@@

```{r}
#| label: fig-study2advice
#| fig-cap: "Advisers were asked to suggest what rank to bet on for someone they ranked as the 7th most attractive out of ten people. When incentivized for accuracy, most of them advised betting on 7, which was based on their evaluation of the recipient's attractiveness. However, when incentivized for likability, many participants gave flattering advice by suggesting betting on higher ranks that imply a higher evaluation of attractiveness."
#| eval: false

@@

```
```{r}
#| label: fig-study2advice
#| fig-cap: "Participants in Study 2 give advice to the person they rank 7th most attractive out of 10. When incentivized for accuracy, the majority recommends betting on that rank. However, when incentivized for likeability, participants inflate their advice to flatter the recipient as more attractive."
study2_time2 %>%
  dplyr::group_by(Condition) %>%
  dplyr::mutate(total_per_condition = dplyr::n()) %>%
  dplyr::group_by(Condition, Advice, total_per_condition) %>%
  dplyr::summarize(n = dplyr::n()) %>%
  dplyr::mutate(pct = ifelse(Condition == "Accurate", -n/total_per_condition, n/total_per_condition),
                Advice = factor(Advice)) %>%
  ggplot(aes(x = pct, y = Advice, fill = Condition)) +
  geom_col() +
  scale_x_continuous("Share of Advice",
                     limits = c(-0.6, 0.6),
                     labels = \(x) scales::percent(abs(x))) +
  scale_y_discrete(limits = rev) +
  # Dashed 'horizontal' line at Advice = 7
  geom_hline(yintercept = 4, col = 'black', linetype = 'dashed', linewidth = 0.7) +
  # Example text annotation (adjust x/y for positioning)
  annotate("text", x = -0.3, y = 9, label = "Accuracy Incentive", hjust = 1, size = 4, fontface = "bold") +
  annotate("text", x = 0.6, y = 9, label = "Likeable Incentive", hjust = 1, size = 4, fontface = "bold") +
  # scale_y_continuous(name = "Advised Rank", breaks = seq(1, 10)) +
  scale_fill_manual(values = c("#2A629A", "#FF7F3E")) +
  papaja::theme_apa() +
  theme(
    legend.position  = "none",
    axis.text.x  = element_text(size = 14),
    axis.text.y  = element_text(size = 14),
    axis.title.x = element_text(size = 16, face = "bold"),
    axis.title.y = element_text(size = 16, face = "bold")
  )
```

@@

Next, we turn our attention to the advisers (see @fig-study2advice). In the
Accuracy condition, those uploading selfies are on average advised to bet on rank
`r round(mean(study2_time2[study2_time2$Condition == "Accurate",]$Advice),2)`.
Notably, this is significantly more attractive than the 7th rank those
advisers had themselves guessed just on the prior screen
(`r papaja::apa_print(t.test(study2_time2[study2_time2$Condition == "Accurate",]$Advice, mu = 7, var.equal = T))$statistic`).
This suggests that even when incentivized for accuracy, participants
offer flattering advice.[^1] Importantly, and as predicted, we find
that advisers in the Likeable treatment recommend betting on a lower
rank, communicating that they think the participant in the selfie is
more attractive
(`r round(mean(study2_time2[study2_time2$Condition == "Likable",]$Advice),2)`,
`r papaja::apa_print(t.test(Advice ~ Condition, data = study2_time2, var.equal = T))$statistic`).
This suggests that advisers infer flattering someone with pleasant advice
would make the adviser appear more likeable. The distribution shown in
@fig-study2advice shows that participants do not simply tell
participants that they are the most attractive person in the group. They
may infer that flattering advise needs to be somewhat realistic to be
believable. We return to this in the general discussion.

[^1]: This shading could be due to concerns of avoiding disappointment.
    However, it could also be that advisers are uncertain about the
    rankings they have given and make a recommendation that combines
    their own belief with a uniform prior. Therefore, our analyses focus
    on the difference between the two conditions.

@@

```{r}
#| label: tbl-study2regs
#| tbl-cap: "When individuals receive advice that implies a high level of attractiveness (lower rank), they tend to perceive the advice giver as more likable (Column 1), warm (Column 2). Column 3 shows that advisors are rated as more trustworthy when they advice lower ranks, but this relationship is only directional."

@@

In Study 3, we replicate the main effect in a different ego-relevant
domain: intelligence. The experimental design is conceptually identical
to Study 1, but this time we utilize a more direct measure of
expectation by asking participants to estimate their scores.

This three-stage experiment begins with measuring actual performance.
Participants are then anchored to low or high performance expectations
and asked to guess how many questions they answered correctly. As in
previous studies, we recruit advisers who are presented with either just
the advisee's true performance or both the true performance and the
number of questions the advisee believes they answered correctly.
Advisers are then asked to provide advice on choosing a competitor.
Finally, we present the advice to the advisees and ask them to make
their final choice. Our key hypotheses are: (1) advisees anchored to
high expectations will be more likely to receive advice to compete
against the high-performing group, and (2) men will be more likely than
women to be advised to compete against the high-performing group when
expectations are disclosed.

@@

We begin by first recruiting a sample of
`r scales::comma(dim(study3_math50)[1])` participants from Prolific to
complete a 10-question multiple choice mathematics quiz. The questions
were taken from a paper-version of the ASVAB standardized exam, such
that answers were not available online. Participants had five minutes to
answer the quiz and were paid 10 cents for each correctly answered
question. Like before, we define top 20 scores as "High Performers" and
the bottom 20 scorers as the "Low Performers." On average, participants
answered `r round(mean(study3_math50$Score),2)` questions correctly,
High Performers scored between `r HP_lower` and `r HP_upper`, and Low
Performers scored between `r LP_lower` and `r LP_upper`. In order to
anchor the expectations of participants in our main experiment, we
simulated 1,000 pairings of groups of five participants, with the
5<sup>th</sup> percentile of groups scoring an average of `r p5` and the
95 th percentile scoring an average of `r p95`. We report these averages
to participants in the Low Expectations and High Expectations treatment,
respectively.

We then recruited `r scales::comma(dim(study3_time1All)[1])`
participants for the role of advisees in Stage 1 of our main experiment.
To arrive at a gender-balanced sample, we dropped the last two male
participants to complete the survey, ending up with a sample of 500 men
and 500 women ($M_{\text{Age}}$ = `r mean(study3_time1$Age)`).
Participants completed the same 10-item mathematics quiz as the earlier
participants and were informed that their performance would affect their
bonus earnings in a follow-up stage to be conducted a few days later.
After completing the quiz, we informed them of the average score of a
group of five participants from the preliminary survey. We randomly
assigned them to learn about the 5<sup>th</sup> percentile of groups,
which scored `r p5`. ("Low Expectations" treatment) or the
95<sup>th</sup> percentile of groups, which scored `r p95` ("High
Expectations" treatment). Participants then made a guess
(unincentivized) about how many questions they think they answered
correctly. The survey concluded with basic demographic questions.

We then recruited `r scales::comma(dim(study3_time2_giver)[1])`
participants for Stage 2, placing them in the role of advisers. We began
by informing them of the mathematics quiz that participants in the
preliminary study and Stage 1 had completed, and informed them of the
average score of all participants in the preliminary study. Advisers had
to recommend whether an advisee should compete against the Low
Performers or the High Performers (we used these terms in the survey).
We anticipated that being told to compete against High Performers is
more flattering and hence being told to compete against the Low
Performers would be disappointing if one had expected to do well.
Advisees could earn a bonus if their score was equal to or higher than
that of a randomly selected member from their chosen group. Competing
against the High Performers was more challenging, resulting in an
average bonus of 50 cents, compared to 30 cents when competing against
other groups.

Like before, advisers were randomly assigned to one of two treatments.
In the "Baseline" treatment, they only observed the score of the advisee
on the mathematics quiz. In the "Expectation" treatment, they observed
the score as well as the advisee's guess for how many questions they
answered correctly. We want to emphasize that sice the outcome is
determined only by the past score on the quiz, the advisee's guess is
immaterial to which group they should compete against. Moreover, in
neither treatment did they receive any demographic information about
their advisee. Participants gave recommendations to 10 advisees, which
unbeknownst to them were five men and five women matched to have
identical performance on the test.[^2] They were informed that if their
advice was shown to a participant who returned for the follow-up survey,
they would receive the identical bonus as that participant. The survey
then concluded with basic demographic questions.

[^2]: We made this decision to account for the possibility of gender
    differences in performance.

Finally, we invited participants from Stage 1 back for the follow-up
survey. Following our preregistration, we kept the survey open for 7
days. In total, `r scales::comma(dim(study3_time3)[1])` participants
(`r scales::comma(dim(study3_time3[study3_time3$Gender == 'Male',])[1])`
men,
`r scales::comma(dim(study3_time3[study3_time3$Gender == 'Female',])[1])`
women) returned. The brief survey reminded them of the task they
completed in Stage 1, informed them that other participants from
Prolific had observed their real score and given them advice against
which group to compete against, and finally were reminded them of how
many questions they guessed they had answered correctly. Importantly,
they were not informed of their true score or the score of the groups
they could compete against. Participants then observed the advice from a
randomly selected adviser and made their decision.

@@

We begin by examining the performance of the Stage 1 participants.
Because our treatment takes place after participants completed the
mathematics quiz, we would not expect a difference in performance across
the Low and High Expectations treatments. Indeed, the two groups scored
no different from one another
(`r round(mean(study3_time1[study3_time1$Condition == "Low",]$count),2)`
and
`r round(mean(study3_time1[study3_time1$Condition == "High",]$count),2)`
for the Low Expectations and High Expectations treatments, respectively,
`r papaja::apa_print(t.test(count ~ Condition, data = study3_time1, var.equal = T))$statistic`).
The expectations treatment did, however, affect how well they thought
they performed. Participants in the "Low Expectations" treatment guessed
a score of
`r round(mean(study3_time1[study3_time1$Condition == "Low",]$count),2)`
vs.
`r round(mean(study3_time1[study3_time1$Condition == "High",]$Guess),2)`
in the High Expectations treatment
(`r papaja::apa_print(t.test(Guess ~ Condition, data = study3_time1, var.equal = T))$statistic`),
showing that the manipulation was successful, albeit small. Contrary to
our expectations, we did find a gender difference in performance: men
scored
`r round(mean(study3_time1[study3_time1$Gender == "Male",]$count),2)` on
average, while women scored
`r round(mean(study3_time1[study3_time1$Gender == "Female",]$count),2)`
(`r papaja::apa_print(t.test(count ~ Gender, data = study3_time1, var.equal = T))$statistic`).
Consistent with this difference, men thought they answered more
questions correctly than did women
(`r round(mean(study3_time1[study3_time1$Gender == "Male",]$Guess),2)`
vs
`r round(mean(study3_time1[study3_time1$Gender == "Female",]$Guess),2)`,
`r papaja::apa_print(t.test(Guess ~ Gender, data = study3_time1, var.equal = T))$statistic`).
This difference, however, does not affect the interpretation of our
findings, which will rely on an interaction of gender with an
experimental treatment for Stage 2 participants.

We define a measure of "overconfidence" as (Performance - Estimate). We
observe that women underestimate their performance by 0.76 points, while
men do so by only 0.27 points
`r papaja::apa_print(t.test(overconfidence ~ Gender, data = study3_time1 %>% dplyr::mutate(overconfidence = Guess - count)))$full_result`.
Although neither gender is overconfident, men on average are more
confident in their performance than are women, as we had expected.

```{r}
#| label: tbl-study3regs
#| tbl-cap: "Column 1 displays the advice given to compete against a High Performance group based on whether the advisees were primed with low or high expectations and the expectation level shown to the advisor. Column 2 displays the advice given to compete against a High Performance group based on the gender of the targets and the expectation level shown to the advisor. Column 3 displays the expected bonus of the advice received based on the gender of the targets and the expectation level shown to the advisor. All standard errors clustered at the level of the advisor."

@@

Next, we examine whether advisers took the advisee's expectations into
account. Column 1 of @tbl-study3regs reports a linear probability model
on advising to compete against the High Performer group for advisers in
the "Expectations" treatment, controlling for the true performance of
the advisee. Because each adviser made ten recommendations, we cluster
standard errors at the adviser level. However, contrary to our
expectations, we do not see a significant effect of the expectations
treatment. This may be because the induced difference in expectations
was too small.

However, recall that men were more confident in their performance than
were women. Our theory thus predicts that showing expectations should
lead to more flattering advice (i.e., a recommendation to compete
against the High Performer group) for men than for women. We report a
linear probability model with advice to compete against the high group
as the outcome measure, and the advisee's gender, whether expectations
were shown to the adviser, and the interaction of the two in Column 2 of
@tbl-study3regs. As predicted, we find a significant interaction effect:
men are more likely to be advised to compete against the High Performers
when expectations are shown than when they are hidden ($p < 0.001$). We
show this result graphically in @fig-study3genderdiff.

>>>>

---
title: "Flattering Advice: Avoiding Disappointment in Advice-Giving"
shorttitle: "Flattering Advice"
date: yes
author: 
  - name: "Amanda Chen"
    corresponding: true
    orcid: 0000-0002-4428-627X
    email: zchengj@connect.ust.hk
    affiliations:
      - id: id1
        name: "The Hong Kong University of Science and Technology"
        department: "Department of Management"
        city: "Hong Kong"
  - name: "David Hagmann"
    orcid: 0000-0002-2080-997X
    affiliations:
    - id: id1
# author-note:
  # disclosures:
    # gratitude: The authors are grateful to XXX for their helpful comments.
abstract: Good advice improves decision quality but often requires delivering unpleasant truths that may disappoint advisees. Across three pre-registered and incentivized experiments involving real adviser-advisee interactions (N = 3,900), we show that advisers prioritize avoiding disappointment at the expense of accuracy and their own earnings. In Study 1, advisers financially rewarded for accuracy still tailor recommendations to aspirational goals expressed by advisees, resulting in worse advice. When incentivized to be liked, advisers provide even more flattering advice, and advisees reward this by rating these advisers as more likable, despite the advice being less honest and less accurate (Study 2). The desire to avoid disappointment may lead to inequities if there are differences in expectations across social groups. In Study 3, we examine a setting in which men expect to perform better than women. We show that advisers take into account these expectations, leading to systematically different advice for men and women even when their gender is unknown to advisers. Advisers' efforts to avoid disappointment may thus contribute to systematic gender disparities in advice, with implications for downstream decision-making.

@@

We present the results of three preregistered experiments in which
participants are paired anonymously as advisers and advisees. Advisers
are incentivized to offer accurate advice, with bonus earnings depending
on the outcome of an advisee's decision. We study a setting in which
advisees can communicate their expectations to an adviser, and
experimentally manipulate whether advisers observe this expectation. Our
design simulates organizational contexts where mentors and managers
often have some sense of an employee's confidence or aspirations and
thus could adjust their feedback to avoid disappointing them.
Interpersonal considerations would likely be stronger in situations in
which the adviser and advisee have an existing relationship and when the
advice is delivered face-to-face. Moreover, advisers usually do not
suffer any direct costs when their flattering advice leads to a poor
outcome for the recipient, particularly when it is not clear what would
have happened under a counterfactual.[^1]

[^1]: In some cases, however, repeated interaction might also increase
    incentives for honest feedback. Someone who is known to persistently
    give overoptimistic advice may be viewed as less trustworthy in the
    long run.

@@

\linespread{1}

```{r}
#| label: tbl-study1regs
#| tbl-cap: "Advice to compete against the High Performer Group in Study 1. Displaying the non-binding choice of those for whom the Low Performer Group was selected by default makes it less likely that they are advised to compete against the High Performer Group (Column 1). Column 2 restricts the analysis to advisers who observe advisees' initial non-binding choice and controls for advisees' score on the quiz. Standard errors are clustered at the adviser level."
#| after-caption-space: 0pt

@@

We now examine whether advisers take the advisees' expectations into
account. Advisers observe the true performance of the advisee, such that
expectations about performance add no additional information. However,
because participants who score higher are also more likely to pick the
High Performer Group, we use our Default treatment as an instrument that
reflect higher expectations. We show linear probability models for the
decision of the Advisers, which is equal to 1 if they recommend
competing against the High Performer Group and 0 if they recommend the
Low Performer Group. In Column 1 of @tbl-study1regs, we show our main
specificiation with the experimental assignment for Advisees (High
Performer Default vs. Low Performer Default) and Advisers (Expectation
Shown vs. Expectation Hidden), as well as their interaction. Because
each adviser made ten recommendations, we cluster standard errors at the
adviser level. We find that displaying expectations of those in the Low
Performer Default reduces recommendations to compete against the Top
group by 4 percentage points, or about 10% relative to when expectations
are hidden. This decrease is marginally smaller for advisees with a
default choice to compete against the High Performer Group (nudged to
have high expectations, though not necessarily holding them), and whose
advice did not differ based on whether their expectations were shown or
hidden. We reported a linear probability model on advising to compete
against the High Performer group for advisers in the "Expectations"
treatment, controlling for the true performance of the advisee in Column
2 of @tbl-study1regs. We observe a significant main effect of the
default treatment, where a high default increased the likelihood of
receiving flattering advice. It is worth noting that, since only
`r round(mean(study1_time1[study1_time1$default == "High",]$Guess),2)*100`
% of advisees in the high default group actually stuck with their
default choice, this represents a conservative test of the effect of
high expectations on the likelihood of receiving flattering advice.

@@

Next, we recruit `r scales::comma(dim(study2_time2)[1])` participants
from Prolific for the role of advisers ($M_{\text{Age}}$ =
`r mean(study2_time2$age)` years;
`r round(mean(study2_time2$gender == "Female")*100, 2)`% Female). We
first collect demographic information, then match them to a group of the
opposite sex. They rank selfies from ten participants from most to least
attractive using a drop-down menu next to each picture. Because of a
limitation with the survey software, we could not validate that each
rank is given only once, and we remove 115 participants who failed to
follow instructions and did not provide a complete ranking.[^2]

[^2]: Moreover, we could not collect data from any participant who did
    not select any participant as the 7th most attractive participant,
    as that image would be shown on subsequent pages.

@@

```{r}
#| label: fig-study2advice-old
#| fig-cap: "Advisers were asked to suggest what rank to bet on for someone they ranked as the 7th most attractive out of ten people. When incentivized for accuracy, most of them advised betting on 7, which was based on their evaluation of the recipient's attractiveness. However, when incentivized for likability, many participants gave flattering advice by suggesting betting on higher ranks that imply a higher evaluation of attractiveness."
#| eval: false

@@

```{r}
#| label: fig-study2advice
#| fig-cap: "Participants in Study 2 give advice to the person they rank 7th most attractive out of 10. When incentivized for accuracy, the majority recommends betting on that rank. However, when incentivized for likeability, participants inflate their advice to flatter the recipient as more attractive."

study2_time2 %>%
  dplyr::group_by(Condition) %>%
  dplyr::mutate(total_per_condition = dplyr::n()) %>%
  dplyr::group_by(Condition, Advice, total_per_condition) %>%
  dplyr::summarize(n = dplyr::n()) %>%
  dplyr::mutate(pct = ifelse(Condition == "Accurate", -n/total_per_condition, n/total_per_condition),
                Advice = factor(Advice)) %>%
  ggplot(aes(x = pct, y = Advice, fill = Condition)) +
  geom_col() +
  scale_x_continuous("Share of Advice",
                     limits = c(-0.6, 0.6),
                     labels = \(x) scales::percent(abs(x))) +
  scale_y_discrete(limits = rev) +
  # Dashed 'horizontal' line at Advice = 7
  geom_hline(yintercept = 4, col = 'black', linetype = 'dashed', linewidth = 0.7) +
  # Example text annotation (adjust x/y for positioning)
  annotate("text", x = -0.3, y = 9, label = "Accuracy Incentive", hjust = 1, size = 4, fontface = "bold") +
  annotate("text", x = 0.6, y = 9, label = "Likeable Incentive", hjust = 1, size = 4, fontface = "bold") +
  # scale_y_continuous(name = "Advised Rank", breaks = seq(1, 10)) +
  scale_fill_manual(values = c("#2A629A", "#FF7F3E")) +
  papaja::theme_apa() +
  theme(
    legend.position  = "none",
    axis.text.x  = element_text(size = 14),
    axis.text.y  = element_text(size = 14),
    axis.title.x = element_text(size = 16, face = "bold"),
    axis.title.y = element_text(size = 16, face = "bold")
  )
```

@@

Next, we turn our attention to the advisers (see @fig-study2advice). In
the Accuracy condition, those uploading selfies are on average advised
to bet on rank
`r round(mean(study2_time2[study2_time2$Condition == "Accurate",]$Advice),2)`.
Notably, this is significantly more attractive than the 7th rank those
advisers had themselves guessed just on the prior screen
(`r papaja::apa_print(t.test(study2_time2[study2_time2$Condition == "Accurate",]$Advice, mu = 7, var.equal = T))$statistic`).
This suggests that even when incentivized for accuracy, participants
offer flattering advice.[^3] Importantly, and as predicted, we find that
advisers in the Likeable treatment recommend betting on a lower rank,
communicating that they think the participant in the selfie is more
attractive
(`r round(mean(study2_time2[study2_time2$Condition == "Likable",]$Advice),2)`,
`r papaja::apa_print(t.test(Advice ~ Condition, data = study2_time2, var.equal = T))$statistic`).
This suggests that advisers infer flattering someone with pleasant
advice would make the adviser appear more likeable. The distribution
shown in @fig-study2advice shows that participants do not simply tell
participants that they are the most attractive person in the group. They
may infer that flattering advise needs to be somewhat realistic to be
believable. We return to this in the general discussion.

[^3]: This shading could be due to concerns of avoiding disappointment.
    However, it could also be that advisers are uncertain about the
    rankings they have given and make a recommendation that combines
    their own belief with a uniform prior. Therefore, our analyses focus
    on the difference between the two conditions.

@@

```{r}
#| label: tbl-study2regs
#| tbl-cap: "When individuals receive advice that implies a high level of attractiveness (lower rank), they tend to perceive the advice giver as more likable (Column 1), warm (Column 2). Column 3 shows that advisors are rated as more trustworthy when they advice lower ranks, but this relationship is only directional."
#| after-caption-space: 0pt

@@

In Study 3, we examine how advisers who want to avoid disappointment may
inadvertently contribute to gender differences in advice. We ask
participants to estimate their score on a mathematics test, and
anticipate that men will guess higher than women. Moreover, we expect
that advisers take into account the expected score when recommending
whether someone should compete against a group of High or Low
performers. As a result, we hypothesize that, when expectations are
known to advisers, men will be advised to compete against High
Performers more than women, even when advisers do not know gender and
expectations are uninformative.

@@

We begin by first recruiting a sample of
`r scales::comma(dim(study3_math50)[1])` participants from Prolific to
complete a 10-question multiple choice mathematics quiz. The questions
are taken from a paper-version of the ASVAB standardized exam, such that
answers are not available online. Participants have five minutes to
answer the quiz and are paid 10 cents for each correctly answered
question. Like in Study 1, we define the top 20 scores as "High
Performers" and the bottom 20 scorers as the "Low Performers." On
average, participants answered `r round(mean(study3_math50$Score),2)`
questions correctly, High Performers scored between `r HP_lower` and
`r HP_upper`, and Low Performers scored between `r LP_lower` and
`r LP_upper`. In order to anchor the expectations of participants in our
main experiment, we simulated 1,000 pairings of groups of five
participants, with the 5<sup>th</sup> percentile of groups scoring an
average of `r p5` and the 95 th percentile scoring an average of
`r p95`. We report these averages to participants in the Low
Expectations and High Expectations treatment, respectively.

We then recruit `r scales::comma(dim(study3_time1All)[1])` participants
for the role of advisees in Stage 1 of our main experiment. To arrive at
a gender-balanced sample, we drop the last two male participants to
complete the survey, ending up with a sample of 500 men and 500 women
($M_{\text{Age}}$ = `r mean(study3_time1$Age)`). Participants complete
the same 10-item mathematics quiz as the earlier participants and are
informed that their performance would affect their bonus earnings in a
follow-up stage to be conducted a few days later. After completing the
quiz, we inform them of the average score of a group of five
participants from the preliminary survey. We randomly assign them to
learn about the 5<sup>th</sup> percentile of groups, which scored
`r p5`. ("Low Expectations" treatment) or the 95<sup>th</sup> percentile
of groups, which scored `r p95` ("High Expectations" treatment).
Participants then make an unincentivized guess about how many questions
they think they answered correctly. The survey concludes with basic
demographic questions.

We then recruit `r scales::comma(dim(study3_time2_giver)[1])`
participants for Stage 2, placing them in the role of advisers. We begin
by informing them of the mathematics quiz that participants in the
preliminary study and Stage 1 have completed, and inform them of the
average score of all participants in the preliminary study. Advisers
have to recommend whether an advisee should compete against the Low
Performers or the High Performers (we use these terms in the survey). We
anticipate that being told to compete against High Performers is more
flattering and hence being told to compete against the Low Performers
would be disappointing if one had expected to do well. Advisees would
earn a bonus if their score is equal to or higher than that of a
randomly selected member from their chosen group. They earn 50 cents if
they outperform a High Performer and 30 cents if they outperform a Low
Performer.

Advisers are randomly assigned to one of two treatments. In the
"Baseline" treatment, they only observe the score of the advisee on the
mathematics quiz. In the "Expectation" treatment, they observe the score
as well as the advisee's guess for how many questions they answered
correctly. Since the outcome is determined only by the past score on the
quiz, the advisee's guess is not informative for the recommendation,
which can be based on the revealed score. Moreover, in neither treatment
did they receive any demographic information about the advisees.
Participants give recommendations to 10 advisees, which unbeknownst to
them are five men and five women matched to have identical performance
on the test.[^4] They are informed that if their advice is shown to a
participant who returns for the follow-up survey, they will receive the
identical bonus as that participant. The survey then concludes with
basic demographic questions.

[^4]: We made this decision to account for the possibility of gender
    differences in performance.

Finally, we invite participants from Stage 1 back for the follow-up
survey. Following our preregistration, we keep the survey open for seven
days. In total, `r scales::comma(dim(study3_time3)[1])` participants
(`r scales::comma(dim(study3_time3[study3_time3$Gender == 'Male',])[1])`
men,
`r scales::comma(dim(study3_time3[study3_time3$Gender == 'Female',])[1])`
women) return. The brief survey reminds them of the task they completed
in Stage 1, informs them that other participants from Prolific have
observed their real score and given them advice on which group to
compete against, and finally are reminded of how many questions they
guessed they had answered correctly. Importantly, they are not informed
of their true score or the score of the groups they could compete
against. Participants then observe the advice from a randomly selected
adviser and make their decision.

@@

We begin by examining the performance of the Stage 1 participants.
Because our treatment takes place after participants complete the
mathematics quiz, we would not expect a difference in performance across
the Low and High Expectations treatments. Indeed, the two groups score
no different from one another
(`r round(mean(study3_time1[study3_time1$Condition == "Low",]$count),2)`
and
`r round(mean(study3_time1[study3_time1$Condition == "High",]$count),2)`
for the Low Expectations and High Expectations treatments, respectively,
`r papaja::apa_print(t.test(count ~ Condition, data = study3_time1, var.equal = T))$statistic`).
As intended, the expectations treatment does affect how well they
thought they performed. Participants in the "Low Expectations" treatment
guess a score of
`r round(mean(study3_time1[study3_time1$Condition == "Low",]$count),2)`
vs.
`r round(mean(study3_time1[study3_time1$Condition == "High",]$Guess),2)`
in the High Expectations treatment
(`r papaja::apa_print(t.test(Guess ~ Condition, data = study3_time1, var.equal = T))$statistic`),
showing that the manipulation is successful, albeit small. Contrary to
our expectations, we find a gender difference in performance: men score
`r round(mean(study3_time1[study3_time1$Gender == "Male",]$count),2)` on
average, while women score
`r round(mean(study3_time1[study3_time1$Gender == "Female",]$count),2)`
(`r papaja::apa_print(t.test(count ~ Gender, data = study3_time1, var.equal = T))$statistic`).
Consistent with this difference, men think they answered more questions
correctly than do women
(`r round(mean(study3_time1[study3_time1$Gender == "Male",]$Guess),2)`
vs
`r round(mean(study3_time1[study3_time1$Gender == "Female",]$Guess),2)`,
`r papaja::apa_print(t.test(Guess ~ Gender, data = study3_time1, var.equal = T))$statistic`).
This difference, however, does not affect the interpretation of our
findings, which will rely on an interaction of gender with an
experimental treatment for Stage 2 participants.

We define a measure of "overconfidence" as (Performance - Estimate). We
observe that women underestimate their performance by 0.76 points, while
men do so by only 0.27 points
(`r papaja::apa_print(t.test(overconfidence ~ Gender, data = study3_time1 %>% dplyr::mutate(overconfidence = Guess - count)))$full_result`).
Although neither gender is overconfident, men on average are more
confident in their performance than are women, as we had expected.

```{r}
#| label: tbl-study3regs
#| tbl-cap: "Column 1 displays the advice given to compete against a High Performance group based on whether the advisees were primed with low or high expectations and the expectation level shown to the advisor. Column 2 displays the advice given to compete against a High Performance group based on the gender of the targets and the expectation level shown to the advisor. Column 3 displays the expected bonus of the advice received based on the gender of the targets and the expectation level shown to the advisor. All standard errors clustered at the level of the advisor."
#| after-caption-space: 0pt

@@

Next, we examine whether advisers take the advisee's expectations into
account. Column 1 of @tbl-study3regs reports a linear probability model
on advising to compete against the High Performer group for advisers in
the "Expectations" treatment, controlling for the true performance of
the advisee. Because each adviser makes ten recommendations, we cluster
standard errors at the adviser level. However, contrary to our
expectations, we do not see a significant effect of the expectations
treatment. This may be because the induced difference in expectations
was too small.

However, recall that men are more confident in their performance than
are women. Our theory thus predicts that showing expectations should
lead to more flattering advice (i.e., a recommendation to compete
against the High Performer group) for men than for women. We report a
linear probability model with advice to compete against the high group
as the outcome measure, and the advisee's gender, whether expectations
were shown to the adviser, and the interaction of the two in Column 2 of
@tbl-study3regs. As predicted, we find a significant interaction effect:
men are more likely to be advised to compete against the High Performers
when expectations are shown than when they are hidden ($p < 0.001$). We
show this result graphically in @fig-study3genderdiff. As the figure
makes clear, this effect is driven by advice given to those who scored
in the middle of the possible range. Thus, when it is clear that someone
should compete against Low or High performers, advisers are not
deferring to expectations.

To examine the quality of advice, we computed the expected bonus
earnings for someone who follows the recommendations. For example, if an
adviser suggests competing against the High Performer group, we match
the advisee against all 20 members of that group and determined how
often their score matches or exceeds that of the member. We then
multiply this number by the respective bonus earnings (50 cents and 30
cents for the High and Low Performer group, respectively). To see if
including expectations leads to worse advise for men, we report a linear
probability model with the experimental treatment of the adviser, the
gender of the advisee, and their interaction in Column 3 of
@tbl-study3regs. Displaying expectations leads men to be advised to
compete against the High Performer group more often, and this advice
turns out to be costly: men receive worse advise than do women when
expectations are displayed, but not in their absence.[^5]
